---
title: "Adobe Analytics / Google Analytics Comparison"
output: html_document
---

```{r setup and data pull, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSiteCatalyst)       # Package for getting the Adobe Analytics data
library(googleAnalyticsR)    # Package for getting the Google Analytics data
library(tidyverse)           # Package(s) for manipulating and visualizing the data
library(plotly)              # For better / interactive charts
library(scales)              # For getting commas in numbers on y-axes

# Configure settings. This requires three values be loaded as R environment variables.
# This is typically set up in a .Renviron file, which would be a file structured as
# follows (without the "#" at the beginning of each line):

# ADOBE_KEY="[username]:[account]"
# ADOBE_SECRET="65f7XXXXXXXXXXXXXXXXXXXXXXXXXX"
# ADOBE_RSID="XXXXXXXXXXXXXX"
# GOOGLE_PROPERTY="XXXXXXXXXX"

# Get the various values that will be used to access the data
aa_key <- Sys.getenv("ADOBE_KEY")
aa_secret <- Sys.getenv("ADOBE_SECRET")
aa_rsid <- Sys.getenv("ADOBE_RSID")
ga_view_id <- Sys.getenv("GOOGLE_VIEWID")

# Set the start and end dates. These can be adjusted to be something other
# than what is currently set below.
start_date <- Sys.Date()-90
end_date <- Sys.Date()-1

# Set the metrics to be compared. This is set up as a data frame where the 
# first column is the way the metric will be labeled. IMPORTANT: Keep these
# values single words (no spaces, no special characters). They won't actually
# get displayed anywhere, and you have to jump through a bunch of additional
# hoops to get nicer label names. The second column is the metric's name in 
# Adobe Analytics, and the third column is the metric's name in Google Analytics.

metrics <- data.frame(
  label = c("UVs","Visits","Pageviews"),
  adobe_analytics = c("uniquevisitors","visits","pageviews"),
  google_analytics = c("users","sessions","pageviews"),
  stringsAsFactors = FALSE)

###########################
# Pull the Adobe Analytics Data
###########################

# Authenticate Adobe Analytics
SCAuth(aa_key, aa_secret)

# Get the data
aa_data <- QueueOvertime(aa_rsid, start_date, end_date, 
                         metrics$adobe_analytics,
                         date.granularity = "day")

# Remove the columns that aren't needed
aa_data <- select(aa_data, -name, -year, -month, -day,
                  -segment.id, -segment.name)

# Change the date column to be Date format rather than POSIXlt
aa_data$datetime <- as.Date(aa_data$datetime)

# Rename the columns to be the label value (this is so it will
# match when we do the same with the GA data)
colnames(aa_data) <- c("Date", metrics$label)

# Add a column that designates the data as being Adobe Analytics data
aa_data$Platform <- "Adobe Analytics"

###########################
# Pull the Google Analytics Data
###########################

# Authenticate Google Analytics. This will need to launch a browser the
# first time it runs, after which it should have the credentials stored. Simply
# load library(googleAnalyticsR) and then run the one line below to do that (usually).
ga_auth()

# Get the data
ga_data <- google_analytics_4(ga_view_id,
                              c(start_date,end_date),
                              metrics = metrics$google_analytics,
                              dimensions = "date",
                              anti_sample = TRUE)

# Rename the columns to be the label value (this is so it will
# match when we do the same with the AA data)
colnames(ga_data) <- c("Date", metrics$label)

# Add a column that designates the data as being Google Analytics data
ga_data$Platform <- "Google Analytics"

###########################
# Combine the two data sets
###########################

# Combine the data for plotting the line charts with the actual values
master_data <- rbind(aa_data, ga_data)

# Create a separate data frame that focuses on the differences between
# the two data sets. 
master_diffs <- full_join(aa_data, ga_data, by = "Date")

# Replace ".x" and ".y" with ".aa" and ".ga" respectively in the column
# names.
colnames(master_diffs) <- gsub("\\.x$", ".aa", colnames(master_diffs)) %>%
                          gsub("\\.y$", ".ga", .)

# Now, cycle through and add columns that calculate the % differences 
# and total differences for each metric. I can't figure out how to do this
# without a loop, but maybe there's a way?
for (i in 1:length(metrics$label)){
  
  # Add a new column that calculates the total difference. This simply
  # takes the current metric value and adds "Total Diff" onto the end of
  # it to make the column name. Then, it grabs both the ".aa" and ".ga"
  # columsn for that metric and subtracts the second column from the first.
  master_diffs[,paste(metrics$label[i],"Total Diff")] <-
    select(master_diffs, contains(metrics$label[i]))[,1] -
    select(master_diffs, contains(metrics$label[i]))[,2]
 
  # Do the same thing, but calculate the % difference 
  master_diffs[,paste(metrics$label[i],"Percent Diff")] <-
    select(master_diffs, contains(metrics$label[i]))[,1] /
    select(master_diffs, contains(metrics$label[i]))[,2] - 1
}


###########################
# Visualization Setup
###########################

# Define theme for the line charts
default_theme <-   theme_bw() +
  theme(axis.text = element_text(size = 10, colour = "grey50"),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.line.x = element_line(colour = "grey30"),
        axis.line.y = element_blank(),
        legend.title = element_blank(),
        legend.background = element_blank(),
        legend.position = "top",
        legend.justification = "center",
        panel.border = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(colour = "grey90"),
        panel.grid.minor = element_blank()
  )

# Set up function to actually plot the line charts

summary_results <- function(metric){
  
  # Pull just the "total diffs" column for the metric. This makes it easier
  # to make the boxplot, as well as to call out summary diff metrics.
  summary_set <- data.frame(Diffs = master_diffs[,paste(metric,"Total Diff")])

  # Create a boxplot to summarize the differences 
  boxplot_gg <- ggplot(summary_set, aes(x = 0, y=Diffs)) + 
    geom_boxplot(fill='#A4A4A4', color="grey50", outlier.colour= "grey70",
                 outlier.fill = "grey70") +
    geom_vline(aes(xintercept=0), color = "grey30", size = 1) +
    guides(fill=FALSE) + coord_flip() + default_theme +
    theme(panel.grid.major.x = element_line(colour = "grey90"),
          panel.grid.major.y = element_blank(),
          axis.text.y = element_blank(),
          axis.line = element_blank())
  
  boxplot_diffs <- ggplotly(boxplot_gg, width = 780, height = 180)
  
  # Create the base chart with actual values for each platform
  actuals_gg <- ggplot(data = master_data,
                       aes_string(x = "Date", y = metric,
                                  color = "Platform")) +
    geom_line() +
    labs(title = "Actual Values") +
    default_theme +
    scale_color_manual(values = c("#009E73", "#D55E00")) +
    scale_y_continuous(labels = comma)
  
  # Use Plotly to make it somewhat interactive
  actuals <- ggplotly(actuals_gg, width = 900, height = 250)
  
  # Create a chart that shows the difference between the
  # values by day as a % of Adobe Analytics.
  percentdiff_gg <- ggplot(data = master_diffs,
                           aes_string(x = "Date", y = paste0("`",metric," Percent Diff`"))) +
    geom_line(color = "grey60") +
    geom_hline(aes(yintercept=0), color = "grey30", size = 0.5) +
    labs(title = "% Difference by Day") +
    default_theme +
    theme(axis.line.x = element_blank(),
          axis.title = element_blank()) +
    scale_color_manual(values = "#DDDDDD") +
    scale_y_continuous(labels = percent)
  
  # Use Plotly to make it somewhat interactive
  percentdiff <- ggplotly(percentdiff_gg, width = 780, height = 180)
  
  # Bundle up all the results to return
  results <- list(boxplot_diffs = boxplot_diffs,
                  daily_diffs = summary_set,
                  linechart_actuals = actuals,
                  linechart_percentdiff = percentdiff)
  
}

```

## Overview

This report compares key metrics from Adobe Analytics to the comparable metrics from Google Analytics for as single site from **`r start_date` to `r end_date`**. For each metric, a boxplot shows the median, quartiles, and outliers of the _daily difference_ between Adobe Analytics and Google Analtyics for the metric. A line chart shows the _actual values reported_ for the metric from both platform. And, another line chart shows the _% difference_ between the results for each day (with the Adobe Analytics value as the basis).

<!-- RMarkdown doesn't readily support loops for output like charts, 
so you need to manually create a section for each metric being included.
That's what's going on from here on out. You just need to tweak the 
value being passed into the summary_results function, and you need
to adjust the r code chunk names so that there aren't any duplicates. -->

--------

### Unique Visitors / Users

```{r uvs base, echo = FALSE, fig.align = "center"}
processed_results <- summary_results("UVs")
```

The mean daily difference between the two platforms was **`r format(round(as.numeric(mean(processed_results$daily_diffs$Diffs)), 1), nsmall=1, big.mark=",")`** unique visitors / users. The median daily difference was **`r format(round(as.numeric(median(processed_results$daily_diffs$Diffs)), 1), nsmall=1, big.mark=",")`** unique visitors / users. The boxplot below shows the median, quartiles, and outliers for the analyzed timeframe.

```{r uvs charts, echo = FALSE, fig.align = "center"}
processed_results$boxplot_diffs
processed_results$linechart_actuals
processed_results$linechart_percentdiff
```

--------

### Visits / Sessions

```{r visits base, echo = FALSE, fig.align = "center"}
processed_results <- summary_results("Visits")
```

The mean daily difference between the two platforms was **`r format(round(as.numeric(mean(processed_results$daily_diffs$Diffs)), 1), nsmall=1, big.mark=",")`** visits / sessions. The median daily difference was **`r format(round(as.numeric(median(processed_results$daily_diffs$Diffs)), 1), nsmall=1, big.mark=",")`** visits / sessions. The boxplot below shows the median, quartiles, and outliers for the analyzed timeframe.

```{r visits charts, echo = FALSE, fig.align = "center"}
processed_results$boxplot_diffs
processed_results$linechart_actuals
processed_results$linechart_percentdiff
```

--------

### Page Views

```{r pageviews base, echo = FALSE, fig.align = "center"}
processed_results <- summary_results("Pageviews")
```

The mean daily difference between the two platforms was **`r format(round(as.numeric(mean(processed_results$daily_diffs$Diffs)), 1), nsmall=1, big.mark=",")`** page views. The median daily difference was **`r format(round(as.numeric(median(processed_results$daily_diffs$Diffs)), 1), nsmall=1, big.mark=",")`** page views. The boxplot below shows the median, quartiles, and outliers for the analyzed timeframe.

```{r pageviews charts, echo = FALSE, fig.align = "center"}
processed_results$boxplot_diffs
processed_results$linechart_actuals
processed_results$linechart_percentdiff
```

